---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)

avocado_data <- read_csv("data/avocado.csv") %>% clean_names()

```

We’ve looked at a few different ways in which we can build models this week, including how to prepare them properly. This weekend we’ll build a multiple linear regression model on a dataset which will need some preparation. The data can be found in the data folder, along with a data dictionary

We want to investigate the avocado dataset, and, in particular, to model the AveragePrice of the avocados. Use the tools we’ve worked with this week in order to prepare your dataset and find appropriate predictors. Once you’ve built your model use the validation techniques discussed on Wednesday to evaluate it. Feel free to focus either on building an explanatory or a predictive model, or both if you are feeling energetic!

As part of the MVP we want you not to just run the code but also have a go at intepreting the results and write your thinking in comments in your script.

Hints and tips

i. region may lead to many dummy variables. Think carefully about whether to include this variable or not (there is no one ‘right’ answer to this!)

ii. Think about whether each variable is categorical or numerical. If categorical, make sure that the variable is represented as a factor.

iii. We will not treat this data as a time series, so Date will not be needed in your models, but can you extract any useful features out of Date before you discard it?

iv. If you want to build a predictive model, consider using either leaps or glmulti to help with this

```{r}
glimpse(avocado_data)
```

**Just a little addition to the data dictionary: it looks like the x number refer to the general size of each of the avocados. So x4046 is "60 or smaller (Small Hass)",  x4225 is "Size 40 & 48 (Large Hass)" and x4770 is "Size 36 & larger (Extra Large Hass)". Think it might be useful to just reassign these in terms of small, medium and large, although will have to figure out if this makes any difference.  

```{r}
avocado_data %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

```{r}
avocado_data %>% 
distinct(year)
```

*Right, so first thoughts should probably be on which variables we need to get rid of. There not the most intuitive of variables it seems but there is certainly some duplication in there.* 

```{r}
#Just checking that there are no duplicates in terms of the dates.

avocado_data %>% 
  get_dupes(date)
```
*Right, so there it looks like there are 108 rows for each date. It also looks like the last of these rows are mislabeled in terms of the x1 label (i.e. the date changes but the x1 number remains the same). Considering the x1 number doesn't seem to have any inherrent value anyway and seems to be mislabeled, this is the first variable that should go.*

*The "year" variable is also an instant contender to go considering the information contained in it is already contained in the "date" variable.*

*The region variable could definitely be useful, but I'm not sure if the observations are consistent or if there is a lot of cleaning to be done in terms of this.*

```{r}
avocado_data %>% 
  distinct(region)
```

*Right, so classic case of not reading the instructions properly. Looks like "date" can be taken out altogether, although we'll probably want to extract the month before we do this.*

```{r}
avocado_data_trim <-
avocado_data %>% 
mutate(month = month.name[month(date)]) %>% 
  select(-date)
```

*Next, I think it's time to decide what to do with the regions variable. I think region's quite an interesting one to explore, but I think I've not left myself too much time to be able to fully work out a plan for this. So what I'm going to do is to try and reduce this list and therefore our data. I will do this in order to access whether the *city *that avocados are bought in has any major effect.*

```{r}
avocado_data_trim <-
avocado_data_trim %>% 
  filter(!region %in% c("California", 
                      "GreatLakes", 
                      "Midsouth", 
                      "Northeast", 
                      "NorthernNewEngland", 
                      "Plains", 
                      "SouthCarolina", 
                      "SouthCentral", 
                      "Southeast", 
                      "TotalUS", 
                      "West", 
                      "WestTexNewMexico")
  ) 

```

```{r}
#Could have done a bit more with this and tidied up the city names a bit.
avocado_data_trim <- 
avocado_data_trim %>% 
  rename(city = region)
```

*Few more bits to tidy up. First let's get rid of that index column at the start. Then let's get every variable that needs to be changed to a factor done. My idea is that we should do the year, month, type and region*

```{r}
avocado_data_trim <-
avocado_data_trim %>% 
  select(-x1) %>% 
  mutate(year = as.factor(year),
         city = as.factor(city),
         type = as.factor(type),
         month = as.factor(month))


```

*Cool, I think that this should work. Time to to have a go at the model building. 