---
title: "R Notebook"
output: html_notebook
---

```{r} 
#Should I automatically load in janitor each times as well and clean the columns?
library(tidyverse)

tweets_codeclan <- read_csv("data/code_clan_tweets.csv")
codeclan_info <- read_csv("data/code_clan_info.csv")
```

```{r}
view(tweets_codeclan)
```


Q1.Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.

```{r}
tweets_codeclan %>%
  dim()

tweets_codeclan %>%
  names()
  
```

Q2. Find the total number of favourited tweets (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).

```{r}
q2_fav_non_quote <-
tweets_codeclan %>% 
  filter(is_quote == "FALSE") %>% 
  summarise(favourite_non_quote = sum(favorite_count))

q2_fav_non_quote
```

Q3. Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.

```{r}

q3_mean_retweets_by_source <-
tweets_codeclan %>%
  group_by(source) %>% 
  filter(is_quote == "FALSE") %>% 
  summarise(retweet_mean = mean(retweet_count))

q3_mean_retweets_by_source

```

```{r}
tweets_codeclan %>% 
  distinct(source)
```

Q4. Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.

```{r}
q4_likes_by_media <-
tweets_codeclan %>%
  mutate(media_type = replace_na(media_type, replace = "text")) %>% 
  group_by(media_type) %>% 
  summarise(favorite_count = n()) %>% 
  arrange(desc(favorite_count))

q4_likes_by_media
```

Q5. Find the mean number of characters that a CodeClan tweet contains.

```{r}
#Gave up on this one. Sorry!
tweet_characters <- "[a-zA-Z0-0!@£$%^&*(),.?/]"
  
combined_tweet_characters <-
tweets_codeclan %>%
  select(text) %>% 
  str_count(tweet_characters)

combined_tweet_characters
```

```{r}
combined_tweet_characters
```


Q6. The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.

```{r}
cc_status_url <- 
  codeclan_info %>% 
  select(tweet_id, status_url)

cc_status_url
```

```{r}
q6_tweets_with_status_url <-
left_join(tweets_codeclan, cc_status_url, by = "tweet_id")

q6_tweets_with_status_url
```

Q7. From your new joined data, create a new tibble codeclan_hashtags containing only the tweet_id and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

```{r}
q7_codeclan_hashtags <-
  q6_tweets_with_status_url %>%
  select(tweet_id, hashtags) %>%
  drop_na() %>%
  mutate(hashtags = str_to_lower(hashtags))


q7_codeclan_hashtags

```

Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.

```{r}
  c_string <- "c\\("

q7_codeclan_hashtags %>% 
  mutate(hashtags_plus = str_detect(hashtags, c_string))
```

