---
title: "R Notebook"
output: html_notebook
---

1. Load the data, clean_names() and explore it.

```{r}
library(prob)
library(janitor)
library(tidyverse)

mobile_data <- read_csv("data/20190928-items.csv") %>%
  clean_names()
```

```{r}
mobile_data %>% 
  head()
```
```{r}
mobile_data %>%
  sample_n(10) %>% 
  select(-c(asin, url, image, review_url))
```


Q2. Find the brand with the highest numbers of phones in the dataset.

```{r}
#Checking that each phone is different
mobile_data %>%
  distinct(asin)
```

```{r}
mobile_data %>% 
  group_by(brand) %>%
  summarise(count = n()) %>% 
  slice_max(count)

```

Q3. For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

Hint - you will need to calculate the mean() and sd() of the ratings to plot the appropriate normal.

```{r}
samsung_data <-
mobile_data %>%
  dplyr::filter(brand == "Samsung")


samsung_rating_stats <-
samsung_data %>%
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating))

samsung_rating_stats
```

```{r}
samsung_data %>%
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 30) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = samsung_rating_stats$mean,
      sd = samsung_rating_stats$sd
    ),
    col = "red"
  )
```
**I'm not quote sure to be honest. It looks like the general trend of the ratings follow a normal distribution quite well. However, it does look like there is maybe a bimodal distribution, with two peaks between 3 and 4. If I had to plunk for an answer though, I'd say yes, relatively normal distribution.**



Q4. We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

```{r}
one_sd_samsung <-
samsung_data %>%
  dplyr::filter(
    rating >= samsung_rating_stats$mean - samsung_rating_stats$sd
    ) %>%
  dplyr::filter(
    rating <= samsung_rating_stats$mean + samsung_rating_stats$sd
    ) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_rating_stats))

two_sd_samsung <-
samsung_data %>%
  dplyr::filter(
    rating >= samsung_rating_stats$mean - sum(2*samsung_rating_stats$sd)
    ) %>%
  dplyr::filter(
    rating <= samsung_rating_stats$mean + sum(2*samsung_rating_stats$sd)
    ) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_rating_stats))

three_sd_samsung <-
samsung_data %>%
  dplyr::filter(
    rating >= samsung_rating_stats$mean - sum(3*samsung_rating_stats$sd)
    ) %>%
  dplyr::filter(
    rating <= samsung_rating_stats$mean + sum(3*samsung_rating_stats$sd)
    ) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_rating_stats))

```


```{r}

sam_one_sd_prop = 
one_sd_samsung$prop_within_1sd/samsung_rating_stats$num
            
sam_two_sd_prop = 
two_sd_samsung$prop_within_2sd/samsung_rating_stats$num
            
sam_three_sd_prop = 
three_sd_samsung$prop_within_2sd/samsung_rating_stats$num

```

```{r}
norm_one_sd_prop = 
pnorm(q = 1) - pnorm(q = -1)

norm_two_sd_prop = 
pnorm(q = 2) - pnorm(q = -2)

norm_three_sd_prop =
pnorm(q = 3) - pnorm(q = -3)
```

```{r}
norm_one_sd_prop - sam_one_sd_prop
```

**Looks like there is a just over 11% more observations within 1 standard deviation of the mean in the case of the samsung ratings when compared to normal distribution**

```{r}
norm_two_sd_prop - sam_two_sd_prop
```

**However, there is are over 3% less observations within two standard deviations of the mean in the case of the samsung ratings when compared to normal distribution**

```{r}
norm_three_sd_prop - sam_three_sd_prop
```

**Finally, it looks like are just under 2% less observations within 3 standard deviations of the mean in the case of the samsung ratings when compared to normal distribution. **


Extensions

```{r}
qqnorm(samsung_data$rating)
```

