---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)

revenue_data <- CodeClanData::qb_revenue_breakdown
monthly_sales <- CodeClanData::qb_monthly_sales
competitors <- CodeClanData::qb_competitors
device_data <- CodeClanData::qb_device_data
```

Q1. Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
col_scheme <- c("#E89FE9", "#50434F", "#B6A7B5", "#F9A472", "#BD7040")

  ggplot(revenue_data) +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = c(
    "Trackers" = "#E89FE9",
    "Smart Watches" = "#50434F",
    "Accessories" = "#B6A7B5",
    "Subscription Services" = "#F9A472"
  )) +
  theme_light()
```

Q2. Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).

```{r}
kwik_bit_data <-
  monthly_sales %>%
  filter(Cashflow == "Kwikbit Unit 1 Sales")

ggplot(kwik_bit_data) +
  aes(x = Date, y = Amount) +
  geom_line() +
  theme_light()
```


Q3. Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
monthly_revenue_and_cost <-
  monthly_sales %>%
  filter(Cashflow == c("Total Revenue", "Cost"))

ggplot(monthly_revenue_and_cost) +
  aes(x = Date, y = Amount, colour = Cashflow) +
  geom_line() +
  theme_light()
```

Q4. Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
ggplot(competitors) +
  aes(x = Year, y = Revenue, fill = Company) +
  geom_area() +
  scale_y_continuous(labels = scales::comma) +
  theme_light() +
  scale_fill_manual(values = c(
    "Competitor 1" = "#E89FE9",
    "Competitor 2" = "#50434F",
    "Competitor 3" = "#B6A7B5",
    "Competitor 4" = "#F9A472"))
```


Q5. Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(competitors) +
  aes(x = Year, y = Revenue, colour = Company) +
  theme_light() +
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  geom_point() + 
    scale_colour_manual(values = c(
    "Competitor 1" = "#E89FE9",
    "Competitor 2" = "#50434F",
    "Competitor 3" = "#B6A7B5",
    "Competitor 4" = "#F9A472"))
  
```

Q6. Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
device_data <-
  device_data %>%
  mutate(time = hms::hms(hours = hours, minutes = mins))


ggplot(device_data) +
  aes(x = time, y = counts, colour = id) +
  geom_line() +
  facet_wrap(~ device) +
  theme_light() +
  scale_fill_manual(values = c(
    "person1" = "#E89FE9",
    "person2" = "#50434F",
    "person3" = "#B6A7B5",
    "person4" = "#F9A472",
    "person5" = "#BD7040"))
  
  


```

