---
title: "R Notebook"
output: html_notebook
---

We are going to use a dataset on house sales in Ames, Iowa over a number of years. We will be interested in the SalePrice of houses (though we should clean_names() the data, so this will end up as sale_price). Load the data and explore it.

```{r}
library(tidyverse)
library(janitor)
library(infer)

sales_data <- read_csv("lab_homework_data/ames.csv") %>% 
  clean_names()
```

```{r}
sales_data %>%
  sample_n(10) %>%
  select(sale_price)

```

Plot a histogram of sale_price. Is the distribution normal?

```{r}
sales_data %>%
  ggplot(aes(x = sale_price)) + 
  geom_histogram(col = "white", fill = "dark blue", bins = 40) +
  labs(x = "mean sale price")
```
**Looks like it's a little bit right-skewed.** 

Calculate the population parameter mean of sale_price. Store it in a variable, as we’ll need this value later when we check if each CI contains the parameter value or not.

```{r}
pop_para_mean <-
  sales_data %>%
  summarise(sales_price_mean = mean(sale_price))

pop_para_mean
```

Now we are going to:

i) take 1000 random samples each of size 100 from sale_price in ames

ii) create a bootstrap sampling distribution of mean(sale_price) from each 100-observation sample (use 1000 reps)

iii) calculate the 95% CI for each bootstrap distribution

iv) store the upper and lower bounds of the CI in a tibble called results

```{r}
sales_sample_results <- tibble()
for (sample_num in 1:1000){

  # get a new 100-observation random sample from ames
  ind_sales_sample <- sales_data %>%
    rep_sample_n(size = 100, reps = 1) %>%
    ungroup() %>%
    select(sale_price)
  
  # create a bootstrap distribution of mean(sale_price)
  # use 1000 reps
  sales_bootstrap <- ind_sales_sample %>%
    specify(response = sale_price) %>%
    generate(reps = 1000, type = "bootstrap") %>%
    calculate(stat = "mean")

  # get 95% CI for this bootstrap distribution
  sales_ci <- sales_bootstrap %>%
    get_ci(level = 0.95, type = "percentile")
  
  # prepare new row for results, storing CI bounds
  ind_sales_result <- c(
    sample_num = sample_num, 
    lower = sales_ci %>% 
      pull(lower_ci), 
    upper = sales_ci %>%
      pull(upper_ci)
  )
  
  # add this_result to growing tibble of results
  
sales_sample_results <- sales_sample_results %>%
    bind_rows(ind_sales_result)

}
```

Pipe results to mutate() to add a new variable popn_param_in_ci which will be TRUE if the population parameter we calculated earlier is between lower and upper, and FALSE otherwise.

```{r}

sales_pop_para_mean <-
sales_sample_results %>%
  mutate(popn_param_in_ci = case_when(
    as.numeric(pop_para_mean) > lower & 
      as.numeric(pop_para_mean) < upper ~ TRUE,
        as.numeric(pop_para_mean) < lower | 
      as.numeric(pop_para_mean) > upper ~ FALSE))

```

```{r}
sales_pop_para_mean %>%
  group_by(popn_param_in_ci) %>% 
  summarise(count = n()/nrow(sales_pop_para_mean))
```

**Shortcut for above in answers**

```{r}
#results <- results %>%
 # mutate(popn_param_in_ci = lower <= param$mean_sale_price & param$mean_sale_price <= upper)

#results %>%
  #summarise(proportion_CI_containing_popn_param = mean(popn_param_in_ci))
```

